---
title: "31 IPL Calibration"
subtitle: "Standard Curve 9/18/20"
author: "Kaitlin Rempfert"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged 
    css: stylesheet.css 
    number_sections: yes 
    toc: yes 
    toc_float: true 
    toc_depth: 3 
    code_folding: show 
editor_options:
  chunk_output_type: console
---


```{r setup, echo = TRUE, message=FALSE, warning=FALSE}
# load required packages

library(MetaboAnalystR)
library(tidyverse)
library(ggplot2)
library(ggpmisc)
library(tools)
library(LOBSTAHS)
library(MSnbase)
library(sqldf)
library(xcms)
library(CAMERA)
library(rsm)
        
#load in functions
source(file.path("scripts", "lipid_scripts.R"))
```


# Load in raw mzXML files (Note: files too big for Github repo; need to download from Metabolights, convert to mzxml, and then load from the folder below)
```{r warning=FALSE}
# specify directories subordinate to the working directory in which the .mzXML files for xcms can be found; per xcms documentation, use subdirectories within these to divide files according to treatment/primary environmental variable and file names to indicate timepoint/secondary environmental variable (e.g., depth)

mzXMLdirs = "data/raw/calibration_9_18_20/mzXML_pos"

# specify which of the directories above you wish to analyze this time through
chosenFileSubset = "data/raw/calibration_9_18_20/mzXML_pos"


mzXMLfiles.raw = list.files(chosenFileSubset, recursive = TRUE, full.names = TRUE)

# provide some feedback to user
print(paste0("Loaded ", length(mzXMLfiles.raw)," mzXML files. These files contain ",subset.polarity," ion mode data. Raw dataset consists of:"))

print(mzXMLfiles.raw)

if (exists("excluded.mzXMLfiles") & length("excluded.mzXMLfiles")>0) {

  excludedfiles = getFNmatches(IDnumlist = excluded.mzXMLfiles, filelist = mzXMLfiles.raw) # index files to be excluded

  print(paste0("The following files will be excluded from processing based on user's input:"))
  print(mzXMLfiles.raw[excludedfiles])

  mzXMLfiles = mzXMLfiles.raw[-excludedfiles] # exclude the files from mzXMLfiles

} else {

  mzXMLfiles = mzXMLfiles.raw

}
```

## Load Phenodata
```{r warning=FALSE}

#make dataframe of phenodata
label <- c ("ng_01", "ng_01", "ng_01", "ng_1","ng_1", "ng_1","ng_25", "ng_25", "ng_25", "ng_5", "ng_5", "ng_5", "ng1", "ng1", "ng1")
standards_pd <- data.frame(sample_name = sub(basename(mzXMLfiles), pattern = ".mzXML", replacement = "", fixed = TRUE), sample_group = label, group_for_grouping = "standards", stringsAsFactors = FALSE)

#read in centroided mzXML files for all runs in standard series
standards_raw <- readMSData(mzXMLfiles, pdata = new("NAnnotatedDataFrame", standards_pd), msLevel. = 1, centroided = TRUE, mode = "onDisk")
```


# Examine Raw Data

## Total Ion Chromatogram
```{r warning=FALSE}
#Plot up total ion chromatograms

#get base peak chromatograms from the raw, centroided data
standards_tis <- chromatogram(standards_raw, aggregationFun = "sum")

#assign colors based on label
group_colors <- paste0(RColorBrewer::brewer.pal(5, "Spectral"))
names(group_colors)<- c("ng_01", "ng_1", "ng_25", "ng_5", "ng1")

#plot base peak chromatograms
plot(standards_tis,  col = group_colors[standards_raw$sample_group])
legend("topright", legend = c("ng_01", "ng_1", "ng_25", "ng_5", "ng1"), fill = group_colors)
```

## Base Peak Chromatograms
```{r warning=FALSE}
#Plot up base peak chromatograms

#get base peak chromatograms from the raw, centroided data
standards_bpis <- chromatogram(standards_raw, aggregationFun = "max")

#plot base peak chromatograms
plot(standards_bpis, col = group_colors[standards_raw$sample_group])
legend("topright", legend = c("blank", "ng_001", "ng_01", "ng_1", "ng_25", "ng_5", "ng1"), fill = group_colors)
```

## Total Ion Current
```{r warning=FALSE}
#Plot up total ion current by file

#extract total ion current for each file
standards_tc <- split(tic(standards_raw), f = fromFile(standards_raw))
#plot
boxplot(standards_tc, ylab = "intensity", col = group_colors[standards_raw$sample_group], main = "Total Ion Current")
legend("topright", legend = c("blank", "ng_001", "ng_01", "ng_1", "ng_25", "ng_5", "ng1"), fill = group_colors)
```

## Plot up raw Extracted Ion Chromatograms (EICs) for each standard
```{r warning=FALSE, cache=TRUE}
#read in database table of masses of most prominent adduct for each standard
db <- readxl::read_xlsx("data/standards/37_IPL_standards.xlsx", sheet = 3)

raw_XIC_plot_standards(standards = db, raw = standards_raw)
```


# Peak picking

## Run CentWave algorithm to pick peaks
```{r warning=FALSE}
#pick peaking with centWave algorithm
cwp1 <- CentWaveParam(
                        ppm = centW.ppm,
                        peakwidth = c(centW.min_peakwidth,centW.max_peakwidth),
                        fitgauss = centW.fitgauss,
                        noise = centW.noise,
                        mzdiff = centW.mzdiff,
                        verboseColumns = centW.verbose.columns,
                        snthresh = centW.snthresh,
                        integrate = centW.integrate,
                        prefilter = centW.prefilter,
                        firstBaselineCheck = TRUE,
                        mzCenterFun = centW.mzCenterFun
)

standards_data <- findChromPeaks(standards_raw, param = cwp, return.type = "XCMSnExp", BPPARAM = bpparam())
```

## Examine Picked Peaks
```{r}
#plot extracted chromatograms of picked peaks for standards
XIC_plot_standards(standards = db, data = standards_data)
```


# Peak Alignment and Grouping
```{r warning=FALSE}
pdp <- PeakDensityParam(
                sampleGroups = standards_data$group_for_grouping,
                bw = density.bw,
                minFraction = density.minfrac,
                minSamples = density.minsamp,
                binSize = density.mzwid,
                maxFeatures = density.max
)

standards_data <- groupChromPeaks(standards_data, param = pdp)

#chromatographic alignment (retention time correction)
pgp <- PeakGroupsParam(
  minFraction = density.minfrac,
  extraPeaks = loess.extra,
  smooth = "loess",
  span = loess.span,
  family = loess.family
)

standards_data<-adjustRtime(standards_data, param = pgp)

#perform grouping again
standards_data <- groupChromPeaks(standards_data, param = pdp)

#visualize alignment 
plotAdjustedRtime(standards_data, col = group_colors[standards_data$sample_group],
                  peakGroupsCol = "grey", peakGroupsPch = 1)

#QC on picked peaks- is intensity normalized?
boxplot(featureValues(standards_data, value="into") +1, 
        log="y", las=2, col = group_colors[standards_data$sample_group], ylab = "intensity")
```


# Fill Peaks 
```{r warning=FALSE}
#fill peaks
fpp <- FillChromPeaksParam(ppm = 5,expandRt = .1)
standards_data = fillChromPeaks(standards_data, param = fpp)
```


# Isotope Annotation in CAMERA
```{r warning=FALSE}
#Applying CAMERA to identify isotopic peaks, create xsAnnotate object, and create CAMERA pseudospectra using correlation of xcms peak groups between and within samples. These pseudospectra are the groups within which the adduct hierarchy and retention time screening criteria will be applied using LOBSTAHS

#need to convert MSnExp to an xset for import into CAMERA; need to manually bring over sample names and groups
standards_xset <- as(standards_data, "xcmsSet")
sampnames(standards_xset) <- pData(standards_data)$sample_name
sampclass(standards_xset) <- pData(standards_data)$group_for_grouping


#fillPeaks on xset to remove NAs for CAMERA
standards_xset2 <- fillPeaks(standards_xset)

# first, a necessary workaround to avoid a import error; see https://support.bioconductor.org/p/69414/
imports = parent.env(getNamespace("CAMERA"
                                  ))
unlockBinding("groups", imports)
imports[["groups"]] = xcms::groups
lockBinding("groups", imports)

# create annotated xset using wrapper annotate(), allowing us to perform all CAMERA tasks at once

standards_xset_a  <- annotate(standards_xset2,

                  quick=FALSE, # set to FALSE because we want to run groupCorr; will also cause CAMERA to run adduct annotation. while LOBSTAHS will do its own adduct identification later, it doesn't hurt to do this now if it lets CAMERA create better pseudospectra
                  sample= NA, # use all samples
                  nSlaves=4, # use 4 sockets

                  # group FWHM settings
                  # using defaults for now

                  sigma=6,
                  perfwhm=0.6,

                  # groupCorr settings
                  # using defaults for now

                  cor_eic_th=0.75,
                  graphMethod="hcs",
                  pval=0.05,
                  calcCiS=TRUE,
                  calcIso=TRUE,
                  calcCaS=FALSE, # weird results with this set to TRUE

                  # findIsotopes settings

                  maxcharge=4,
                  maxiso=4,
                  minfrac=0.25, # 0.25?

                  # adduct annotation settings

                  psg_list=NULL,
                  rules=NULL,
                  polarity=subset.polarity,
                  multiplier=3,
                  max_peaks=100,

                  # common to multiple tasks

                  intval="into",
                  ppm=2.5,
                  mzabs=0.0015

                  )

cleanParallel(standards_xset_a) # kill sockets

#remove NAs from groupInfo
#xset_a@groupInfo[is.na(xset_a@groupInfo)] <- 0

print(standards_xset_a)
saveRDS(standards_xset_a, "output/calibration/9_18_20_37IPLs_xset_a.RDS")
```


# LOBSTAHS database screening
```{r warning=FALSE}
LOBdb_31std = generateLOBdbase(polarity = "positive", gen.csv = TRUE, component.defs = "LOBSTAHS_db_generation/31_IPL_stds/LOBSTAHS_componentCompTable_31stds.csv", AIH.defs = "LOBSTAHS_db_generation/31_IPL_stds/LOBSTAHS_adductHierarchies_31stds.csv",  acyl.ranges = NULL, oxy.ranges = NULL)

standards_xset_a <- readRDS("data/calibration/9_18_20_37IPLs_xset_a.RDS")

#load in LOBSTAHS database
LOBdbase_31std <- loadLOBdbase("LOBSTAHS_db_generation/31_IPL_stds/LOBSTAHS_lipid-oxy_DB_pos_2023-03-30T13-28-03-0700.csv", polarity = "positive")

#screen annotated dataset against 32 std database
LOBset_31std <- doLOBscreen(standards_xset_a, polarity = "positive", database = LOBdbase_31std, remove.iso = TRUE, rt.restrict = TRUE, rt.windows = "LOBSTAHS_db_generation/31_IPL_stds/LOBSTAHS_rt.windows_31stds.csv", exclude.oddFA = FALSE, match.ppm = 5, retain.unidentified = FALSE)

#diagnostics of screening;35 IDs
LOBscreen_diagnostics(LOBset_31std)

#get peaklist
peaklist <- getLOBpeaklist(LOBset_31std, include.iso = TRUE, include.unidentified = TRUE, gen.csv = FALSE)

as_tibble(peaklist)
```


# Concentration Calibration Curves
```{r warning= FALSE}
#manually remove duplicated peak assignments by retention time 
peaklist_filt <- peaklist %>% filter(!(match_ID %in% c("8", "30", "36", "37", "38")))

#select necessary columns from LOBSTAHS peaklist
positions <- c(2, 13:27)
standards <- peaklist_filt %>% select(positions) 

#summarize (sum multiple peaks of same standard)
standards_summary <- standards %>% group_by(compound_name) %>% summarize_each(funs(sum))
#transpose 
conc <- standards_summary %>% select(-compound_name) %>% t() %>% as.data.frame()
#make standard names the column names
colnames(conc) <- standards_summary$compound_name
#make sample names a column instead of the rownames
conc <- conc %>% rownames_to_column(var = "id")
#pull out QEF number and concentrations from id
conc <- conc %>% mutate(
  QEF = str_extract(id, ".+?(?=_)"),
  conc = str_extract(id, "(?<=37_).+?(?=ng)")
)
#delete id column
conc$conc <- as.numeric(sub("_", ".", conc$conc))
conc <- conc %>% select(-id) 
#10 uL std injected; convert conc in ng/uL to ng on column
conc <- conc %>% mutate(
  conc = conc * 10
) 

#reshape dataframe for plotting
conc_reshape <- reshape2::melt(conc, id = c("QEF", "conc"), variable.name = "standard", value.name = "area")  
conc_reshape$area <- as.numeric(conc_reshape$area)
conc_reshape$conc <- as.numeric(conc_reshape$conc)

#replace 0's from filling to NAs
conc_reshape <- conc_reshape %>% mutate_if(is.numeric, funs(ifelse(. == 0, NA, .)))
#conc_reshape <- conc_reshape %>% filter(area != "NA")


#plot standard curves with second order polynomial regressions for each standard
conc_plot <- conc_reshape %>% ggplot(
  aes(x= conc, y= area, color = as.factor(conc), group = standard)) +
    geom_point() +
    facet_wrap(vars(standard), scales = "free_y") +
    geom_smooth(aes(linetype = "linear fit"), method = "lm", formula = y ~ x + 0, se = FALSE, show.legend = TRUE) +
   stat_poly_eq(formula = y ~ x + 0, 
                aes(label = paste(..eq.label.., sep = "~~~")), size = rel(2.5), parse = TRUE, label.y = 1.0) +
   stat_poly_eq(formula = y ~ x + 0, 
                aes(label = paste(..rr.label.., sep = "~~~")), size = rel(2.5), parse = TRUE, label.y = .9) +
    geom_hline(aes(linetype = "area minimum for sample processing", yintercept = 50000)) +
    scale_x_continuous() +
    scale_y_continuous() +
    theme(axis.text.x = element_text(angle=90, hjust=1)) +
    labs(color = "ng standard on column", title = "Calibration Curves for IPL standards", x= "ng standard on column", y= "integrated peak area", linetype = NULL)
conc_plot

#calculate models for each standard 
linear_models_9_18_20 <- conc_reshape %>% 
  select(-QEF) %>% 
  group_by(standard) %>%
  nest() %>%
  mutate(
    fit = map(data, ~lm(area ~ conc + 0, data = .x)),
    summary = map(fit, broom::glance),
    tidied = map(fit, broom::tidy)) %>%
  unnest(tidied)

saveRDS(linear_models_9_18_20, "output/calibration/response_factors_9_18_20.RDS")
#need to use package investr to use inverse estimation?
```
