---
title: "Samail Fluid IPLs annotation"
author: "Kaitlin Rempfert"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged 
    number_sections: yes 
    css: stylesheet.css 
    toc: yes 
    toc_float: true 
    toc_depth: 3 
    code_folding: show 
editor_options:
  chunk_output_type: console
---

```{r setup, echo = TRUE, message=FALSE, warning=FALSE}
# load required packages

library(MetaboAnalystR)
library(tidyverse)
library(ggplot2)
library(ggpmisc)
library(tools)
library(LOBSTAHS)
library(MSnbase)
library(sqldf)
library(xcms)
library(CAMERA)
library(rsm)
library(Hmisc)
library(data.table)

#load in functions
source("scripts/lipid_scripts.R")     
```


# Load in raw mzXML files (Note: files too big for Github repo; need to download from Metabolights, convert to mzxml, and then load from the folder below)
```{r warning=FALSE, eval = FALSE}
#path to .mzXML files
dda_folder_fluids_woBA1A = "data/raw/fluids_woBA1A"
#list all files inside specified folder
mzXMLfiles_fluids_woBA1A <- list.files(dda_folder_fluids_woBA1A, recursive = TRUE, full.names = TRUE)

#make phenodata dataframe with grouping labels
label <- c("blank","sample", "sample", "sample", "sample", "blank", "sample", "sample")
pd_fluids_woBA1A <- data.frame(sample_name = sub(basename(mzXMLfiles_fluids_woBA1A), pattern = ".mzXML", replacement = "", fixed = TRUE), sample_group = label, group_for_grouping = "fluids_woBA1A", stringsAsFactors = FALSE)

#load data with phenodata
dda_data_fluids_woBA1A <- readMSData(mzXMLfiles_fluids_woBA1A, pdata = new("NAnnotatedDataFrame", pd_fluids_woBA1A), centroided = TRUE, mode = "onDisk")

#check # of peaks in each MS level to ensure files were converted properly
table(msLevel(dda_data_fluids_woBA1A))

#make sure that precursor intensities are not NA
dda_data_fluids_woBA1A %>%
    filterMsLevel(2L) %>%
    precursorMz() %>%
    head()

#separate positive and negative scans
dda_data_fluids_woBA1A_neg <- dda_data_fluids_woBA1A %>% filterPolarity(polarity. = 0)
dda_data_fluids_woBA1A_pos <- dda_data_fluids_woBA1A %>% filterPolarity(polarity. = 1)
```


## Total Ion Chromatograms
```{r warning=FALSE, eval = FALSE}
#get total ion chromatograms from the raw, centroided data
tis <- chromatogram(dda_data_fluids_woBA1A_pos, aggregationFun = "sum")

#assign colors based on label
group_colors <- paste0(RColorBrewer::brewer.pal(3, "Dark2"))
names(group_colors)<- c("blank", "sample")

#plot total ion chromatograms
plot(tis,  col = group_colors[dda_data_fluids_woBA1A_pos$sample_group], main = "Total Ion Chromatogram")
legend("topright", legend = c("blank", "sample"), fill = group_colors) 
```

## Base Peak Chromatograms
```{r warning=FALSE, eval = FALSE}
#get base peak chromatograms from the raw, centroided data
bpis <- chromatogram(dda_data_fluids_woBA1A_pos, aggregationFun = "max")

#plot base peak chromatograms
plot(bpis, col = group_colors[dda_data_fluids_woBA1A_pos$sample_group], main = "Base Peak Chromatogram")
legend("topright", legend = c("blank", "sample"), fill = group_colors)
```

## Total Ion Current
```{r warning=FALSE, eval = FALSE}
#Plot up total ion current by file

#extract total ion current for each file
tc <- split(tic(dda_data_fluids_woBA1A_pos), f = fromFile(dda_data_fluids_woBA1A_pos))
#plot
boxplot(tc, ylab = "intensity", col = group_colors[dda_data_fluids_woBA1A_pos$sample_group], main = "Total Ion Current")
legend("topright", legend = c("blank", "sample"), fill = group_colors)
```

## Plot up raw Extracted Ion Chromatograms (EICs) for internal standards
```{r warning=FALSE, cache=TRUE, eval = FALSE}
#read in database table of masses of most prominent adduct for each compound
standards <- readxl::read_xlsx("data/standards/internal_standards.xlsx", sheet = 1)  
standards

#function to plot raw EICs
raw_XIC_plot <- function(db, raw) {
  group_colors <- paste0(RColorBrewer::brewer.pal(3, "Dark2"))
  names(group_colors)<- c("blank", "sample")
  
  for (row in 1:nrow(db)){
    XIC <- xcms::chromatogram(raw, mz = c(db$mz_db[row]-.005, db$mz_db[row]+.005), rt = c(db$rt_db[row]*60 - 240, db$rt_db[row]*60 + 240)) 
    plot(XIC, col = group_colors[raw$sample_group], main = db$standard[row])
    legend("topright", legend = c("blank", "sample"), fill = group_colors)
  }
}

#plot
raw_XIC_plot(db = standards, raw = dda_data_fluids_woBA1A_pos)
```


# Pick Peaks
```{r eval=FALSE}
#peak picking parameters
cwp1 <- CentWaveParam(
                        ppm = centW.ppm,
                        peakwidth = c(centW.min_peakwidth,centW.max_peakwidth),
                        fitgauss = centW.fitgauss,
                        noise = centW.noise,
                        mzdiff = centW.mzdiff,
                        verboseColumns = centW.verbose.columns,
                        snthresh = centW.snthresh,
                        integrate = centW.integrate,
                        prefilter = centW.prefilter,
                        firstBaselineCheck = TRUE,
                        mzCenterFun = centW.mzCenterFun
)


#pick peaks
dda_data_fluids_woBA1A_peaks_pos_pp1 <- findChromPeaks(dda_data_fluids_woBA1A_pos, param = cwp1)
dda_data_fluids_woBA1A_peaks_neg_pp1 <- findChromPeaks(dda_data_fluids_woBA1A_neg, param = cwp1)
```


## Plot up EICs of picked peaks
```{r warning=FALSE, eval = FALSE}
#XIC plotting of peaks
XIC_plot <- function(db, data) {
  group_colors <- paste0(RColorBrewer::brewer.pal(3, "Dark2"))
  names(group_colors)<-  c("blank", "sample")
  
  for (row in 1:nrow(db)){
    XIC <- xcms::chromatogram(data, mz = c(db$mz_db[row]-.005, db$mz_db[row]+.005), rt = c(db$rt_db[row]*60 - 240, db$rt_db[row]*60 + 240)) 
    sample_colors <- group_colors[XIC$sample_group]
    plot(XIC, col = sample_colors[chromPeaks(XIC)[, "sample"]], lwd = 2, peakBg = sample_colors[chromPeaks(XIC)[, "sample"]], main = db$standard[row])
    legend("topright", legend = c("blank", "sample"), fill = group_colors)
  }
}

#plot extracted chromatograms of picked peaks for standards
XIC_plot(db = standards, data = dda_data_fluids_woBA1A_peaks_pos_pp1)
```


# Group Peaks and Correct Retention Time 
```{r eval = FALSE, warning = FALSE}
#peak grouping parameters
pdp1 <- PeakDensityParam(
                sampleGroups = dda_data_fluids_woBA1A$group_for_grouping,
                bw = density.bw,
                minFraction = .125,
                minSamples = density.minsamp,
                binSize = density.mzwid,
                maxFeatures = density.max
)

pgp1 <- PeakGroupsParam(
  minFraction = density.minfrac,
  extraPeaks =loess.extra,
  smooth = "loess",
  span = loess.span,
  family = loess.family
)

#peak correlation and retention time correction 
#pos pp1
dda_data_fluids_woBA1A_grouped_pos_pp1 <- groupChromPeaks(dda_data_fluids_woBA1A_peaks_pos_pp1, param = pdp1)
dda_data_fluids_woBA1A_rt_pos_pp1  <- adjustRtime(dda_data_fluids_woBA1A_grouped_pos_pp1, param = pgp1)
dda_data_fluids_woBA1A_rt_grouped_pos_pp1  <- groupChromPeaks(dda_data_fluids_woBA1A_rt_pos_pp1, param = pdp1)
#neg pp1
dda_data_fluids_woBA1A_grouped_neg_pp1 <- groupChromPeaks(dda_data_fluids_woBA1A_peaks_neg_pp1, param = pdp1)
dda_data_fluids_woBA1A_rt_neg_pp1  <-adjustRtime(dda_data_fluids_woBA1A_grouped_neg_pp1, param = pgp1)
dda_data_fluids_woBA1A_rt_grouped_neg_pp1  <- groupChromPeaks(dda_data_fluids_woBA1A_rt_neg_pp1, param = pdp1)

#filter retention time to exclude features that elute in first 1.5 minutes
dda_data_fluids_woBA1A_rt_grouped_rtfilt_pos_pp1 <- dda_data_fluids_woBA1A_rt_grouped_pos_pp1 %>% filterRt(rt= c(150, 1500))
dda_data_fluids_woBA1A_rtfilt_grouped_neg_pp1 <- dda_data_fluids_woBA1A_rt_grouped_neg_pp1 %>% filterRt(rt= c(150, 1500))
```

## Visualize Grouping and Retention Time Adjustments
```{r warning=FALSE, eval = FALSE}
#visualize alignment 
plotAdjustedRtime(dda_data_fluids_woBA1A_rt_grouped_rtfilt_pos_pp1, col = group_colors[dda_data_fluids_woBA1A_rt_grouped_rtfilt_pos_pp1$sample_group],
                  peakGroupsCol = "grey", peakGroupsPch = 1)
legend("bottomright", legend = c("blank", "sample"), fill = group_colors)

#QC on picked peaks- is intensity normalized?
boxplot(featureValues(dda_data_fluids_woBA1A_rt_grouped_rtfilt_pos_pp1, value="into") +1, 
        log="y", las=2, col = group_colors[dda_data_fluids_woBA1A_rt_grouped_rtfilt_pos_pp1$sample_group], ylab = "intensity", main = "Peak Intensities")
legend("topright", inset=c(-0.2,0), legend = c("blank", "sample"), fill = group_colors)
```


# Fill Peaks
```{r eval = FALSE}
dda_data_fluids_woBA1A_rt_grouped_rtfilt_pos_pp1 <- readRDS("fluids_woBA1A_pp1_pos_grouped_rtfilt.rds")
#filter to just MS1 for positive
dda_data_fluids_woBA1A_rt_grouped_rtfilt_pos_pp1_ms1 <- dda_data_fluids_woBA1A_rt_grouped_rtfilt_pos_pp1 %>% filterMsLevel(msLevel. = 1)
#fill peaks
fpp <- FillChromPeaksParam(ppm = 5, expandRt = .1)
fluids_woBA1A_fill = fillChromPeaks(dda_data_fluids_woBA1A_rt_grouped_rtfilt_pos_pp1_ms1, param = fpp)
```


```{r eval = FALSE}
#Applying CAMERA to identify isotopic peaks, create xsAnnotate object, and create CAMERA pseudospectra using correlation of xcms peak groups between and within samples. These pseudospectra are the groups within which the adduct hierarchy and retention time screening criteria will be applied using LOBSTAHS

#need to convert MSnExp to an xset for import into CAMERA 
fluids_woBA1A_ddms2_xset <- as(fluids_woBA1A_fill, "xcmsSet")
#need to manually bring over sample names and groups
sampnames(fluids_woBA1A_ddms2_xset) <- pData(dda_data_fluids_woBA1A)$sample_name
sampclass(fluids_woBA1A_ddms2_xset) <- pData(dda_data_fluids_woBA1A)$group_for_grouping


#fillPeaks on xset to remove NAs for CAMERA (changing NAs to 0 within xset structure)
fluids_woBA1A_ddms2_xset <- fillPeaks(fluids_woBA1A_ddms2_xset)

# first, a necessary workaround to avoid a import error; see https://support.bioconductor.org/p/69414/
imports = parent.env(getNamespace("CAMERA"
                                  ))
unlockBinding("groups", imports)
imports[["groups"]] = xcms::groups
lockBinding("groups", imports)

# create annotated xset using wrapper annotate(), allowing us to perform all CAMERA tasks at once
fluids_woBA1A_ddms2_xset_a <- CAMERA::annotate(fluids_woBA1A_ddms2_xset,

                  quick=FALSE, # set to FALSE because we want to run groupCorr; will also cause CAMERA to run adduct annotation. while LOBSTAHS will do its own adduct identification later, it doesn't hurt to do this now if it lets CAMERA create better pseudospectra
                  sample= NA, # use all samples
                  nSlaves=4, # use 4 sockets

                  # group FWHM settings
                  # using defaults for now

                  sigma=6,
                  perfwhm=0.6,

                  # groupCorr settings
                  # using defaults for now

                  cor_eic_th=0.75,
                  graphMethod="hcs",
                  pval=0.05,
                  calcCiS=TRUE,
                  calcIso=TRUE,
                  calcCaS=FALSE, # weird results with this set to TRUE

                  # findIsotopes settings

                  maxcharge=4,
                  maxiso=4,
                  minfrac=0.5, # 0.25?

                  # adduct annotation settings

                  psg_list=NULL,
                  rules=NULL,
                  polarity=subset.polarity,
                  multiplier=3,
                  max_peaks=100,

                  # common to multiple tasks

                  intval="into",
                  ppm=2.5,
                  mzabs=0.0015

                  )

cleanParallel(fluids_woBA1A_ddms2_xset_a) # kill sockets


print(fluids_woBA1A_ddms2_xset_a)
saveRDS(fluids_woBA1A_ddms2_xset_a, "output/processing/fluids_woBA1A_ddms2_xset_a.RDS")
```


# Identification of Peaks Using LOBSTAHS
```{r eval = FALSE}
fluids_woBA1A_ddms2_xset_a <- readRDS("output/processing/fluids_woBA1A_ddms2_xset_a.RDS")

#load in LOBSTAHS database
LOBdbase_updated <- loadLOBdbase("LOBSTAHS_db_generation/Full_database/LOBSTAHS_lipid-oxy_DB_pos_2021-01-13T11-44-48_PM-0700_updated_1_22_21.csv", polarity = "positive")

#screen annotated dataset against database
LOBset_fluids_woBA1A_ddms2_2 <- doLOBscreen(fluids_woBA1A_ddms2_xset_a, polarity = "positive", database = LOBdbase_updated, remove.iso = TRUE, rt.restrict = TRUE, rt.windows = "LOBSTAHS_db_generation/Full_database/LOBSTAHS_rt.windows_arch.csv", exclude.oddFA = FALSE, match.ppm = 15, retain.unidentified = TRUE)

#diagnostics of screening
LOBscreen_diagnostics(LOBset_fluids_woBA1A_ddms2_2)

#get peaklist
fluids_woBA1A_ddms2_peaklist <- getLOBpeaklist(LOBset_fluids_woBA1A_ddms2_2, include.iso = TRUE, include.unidentified = TRUE, gen.csv = TRUE)
saveRDS(fluids_woBA1A_ddms2_peaklist, "output/processing/fluids_woBA1A_ddms2_peaklist.rds")
```


# MS2 Peak Extraction 
```{r}
#positive
#extract MS2 spectra for features (peak groups) 
dda_spectra_fluids_woBA1A_pos_pp1 <- featureSpectra(dda_data_fluids_woBA1A_rt_grouped_rtfilt_pos_pp1)
#clean spectra
dda_spectra_fluids_woBA1A_pos_pp1 <- clean(dda_spectra_fluids_woBA1A_pos_pp1, all = TRUE)
#combine spectra for same feature_id
dda_combined_fluids_woBA1A_pos_pp1 <- combineSpectra(dda_spectra_fluids_woBA1A_pos_pp1, fcol = "feature_id", method = consensusSpectrum, mzd = 0.005, minProp = 0.05, weighted = FALSE,intensityFun = median, mzFun = median)

#negative
dda_spectra_fluids_woBA1A_neg_pp1 <- featureSpectra(dda_data_fluids_woBA1A_rt_grouped_rtfilt_neg_pp1)
dda_spectra_fluids_woBA1A_neg_pp1 <- clean(dda_spectra_fluids_woBA1A_neg_pp1, all = TRUE)
dda_combined_fluids_woBA1A_neg_pp1 <- combineSpectra(dda_spectra_fluids_woBA1A_neg_pp1, fcol = "feature_id", method = consensusSpectrum, mzd = 0.005, minProp = 0.05, weighted = FALSE,intensityFun = median, mzFun = median)
```


# MS2 verification of LOBSTAHS identifications 

## Load Relevant Databases
```{r}
#load relevant databases
fragments_pos <- read.csv("data/fragment_databases/Fragments_pos_v3.csv")
nl_pos <- read.csv("data/fragment_databses/Neutral_losses_pos_v5.csv")
background_pos <- read.csv("data/fragment_databases/background_pos.csv")

fragments_neg <- read.csv("data/fragment_databases/Fragments_neg.csv")
background_neg <- read.csv("data/fragment_databases/background_neg.csv")
nl_neg <- read.csv("data/fragment_databases/Neutral_losses_neg.csv")

#load relevant databases
fragments_pos_core <- read.csv("data/fragment_databases/Fragments_pos_core.csv")
nl_pos_core <- read.csv("data/fragment_databases/Neutral_losses_pos_core.csv")
background_pos_core <- read.csv("data/fragment_databases/background_pos_core.csv")
```

## Plot annotated spectra for manual inspection 
```{r}
plot_spectra_annotate(spectra = dda_combined_fluids_woBA1A_pos_pp1, fragment_db = fragments_pos, nl_db = nl_pos, background_db = background_pos, to_include = NULL)
```

## Annotate peaklist
```{r eval=FALSE}
#load peaklist
#load .csv files with headgroup information for each lipid species and metadata
fluids_woBA1A_ddms2_peaklist <- readRDS("output/processing/fluids_woBA1A_ddms2_peaklist.rds") 
hg <- read.csv("data/annotation/LOBSTAHS_hg.csv")
hg_meta <- read.csv("data/annotation/hg_meta_3.csv")

#join hg metadata with peaklist
peaklist_w_hg <- left_join(fluids_woBA1A_ddms2_peaklist, hg, by = "species")
peaklist_w_hg_meta <- left_join(peaklist_w_hg, hg_meta, by = "hg_response")
peaklist_w_hg_meta <- peaklist_w_hg_meta %>% rownames_to_column("match_ID2")

#hg annotation of spectra
ms2_features_hg_pos  <- ms2_annotate(spectra = dda_combined_fluids_woBA1A_pos_pp1, fragment_db = fragments_pos, nl_db = nl_pos, background_db = background_pos, intensity_filter = .1)

#note, CANNOT join by feature_id because MS2 feature_id does not correlate with MS1; instead join by mass and retention time range
#extract peaklist columns for joining
peaklist_dt <- peaklist_w_hg_meta %>% select(c("peakgroup_mz", "peakgroup_rt", "match_ID2", "xcms_peakgroup"))
#extract annotation columns for joining
ms2_dt <- ms2_features_hg_pos 

#join with sql allowing multiple matches
id1 <- sqldf("select *
from peaklist_dt peak
left join ms2_dt ms2 on abs(ms2.ms2_precursor - peak.peakgroup_mz) < 0.05
	and abs(ms2.ms2_rt - peak.peakgroup_rt) < 15.0")

#filter ms2 id table to only include unique ms2 ids per xcms peakgroup
id1_filt <- id1 %>% filter(!is.na(ms2_precursor)) %>% select(-c("match_ID2", "peakgroup_rt", "peakgroup_mz", "mtc")) %>% unique() %>% dplyr::rename(id_hg = id) 

#join ms2 data with peaklist and filter to include only compounds that have ms2 data
peaklist_hg <- left_join(peaklist_w_hg_meta, id1_filt, by = "xcms_peakgroup") %>% filter(!is.na(id_hg))
#calculate which peaks have areas at least ten times greater than the blank average
peaklist_hg_blank <- peaklist_hg %>% dplyr::rename(blank_1 = QEF6213_KRR_IPLs_DCM_MeOH_blank_Sep2020opt_dual_ddMS2, blank_2 = QEF6218_KRR_IPLs_DCM_MeOH_blank_Sep2020opt_dual_ddMS2) %>% 
  mutate(
  blank = rowMeans(select(., starts_with("blank")), na.rm = TRUE)) %>% rowwise() %>% mutate(background = ifelse(any(c_across(starts_with("QEF")) > 10*blank), NA, "background")) %>% ungroup()
  
#filter out background peaks and select only peaks that have a hg from LOBSTAHS that matches a hg id from MS2
peaklist_hg_backgroundremoved_match <- peaklist_hg_blank %>% filter(is.na(background)) %>% filter(hg == id_hg)
#peaks that have a MS2 and are not background, but do not match compound in database
peaklist_hg_backgroundremoved_nomatch <- peaklist_hg_blank %>% filter(is.na(background)) %>% filter(is.na(hg))
#add negative precursor mass for peaks
peaklist_hg_backgroundremoved_match <- peaklist_hg_backgroundremoved_match  %>% mutate(neg_precursor_mass = case_when(neg_adduct == "formate" ~ as.numeric(LOBdbase_exact_parent_neutral_mass) + 44.99765, neg_adduct == "hydrogen" ~ as.numeric(LOBdbase_exact_parent_neutral_mass) - 1.00783, TRUE ~ NA_real_))

#negative mode core annotation
ms2_features_neg  <- ms2_annotate(spectra = dda_combined_fluids_woBA1A_neg_pp1, fragment_db = fragments_neg, nl_db = nl_neg, background_db = background_neg, intensity_filter = .1)

#prepare peaklist for matching with negative mode annotation
peaklist_dt2 <- peaklist_hg_backgroundremoved_match %>% select(c("neg_precursor_mass", "peakgroup_rt", "match_ID2"))
ms2_dt2 <- ms2_features_neg %>% dplyr::rename(id_neg = id)

id2 <- sqldf("select *
from peaklist_dt2 peak
left join ms2_dt2 ms2 on abs(ms2.ms2_precursor - peak.neg_precursor_mass) < 0.05
	and abs(ms2.ms2_rt - peak.peakgroup_rt) < 15.0")

#filter ms2 id table to only include unique ms2 ids per xcms peakgroup
id2_filt <- id2 %>% filter(!is.na(ms2_precursor) & !is.na(id_neg)) %>% select(-c("match_ID2")) %>% unique() %>% select(-c(peakgroup_rt, ms2_rt, ms2_precursor, ms2_feature_id)) %>% dplyr::rename(mtc_neg_core_frag = mtc)

#rejoin with peaklist
peaklist_hg_backgroundremoved_match_neg <- left_join(peaklist_hg_backgroundremoved_match, id2_filt, by = "neg_precursor_mass") 

#ms2 positive mode annotation of cores
ms2_features_pos_core <- ms2_annotate(spectra = dda_combined_fluids_woBA1A_pos_pp1, fragment_db = fragments_pos_core, nl_db = nl_pos_core, background_db = background_pos_core, intensity_filter = .1)

#prepare peaklist for matching with positive mode core annotation
peaklist_dt3 <- peaklist_hg_backgroundremoved_match_neg %>% select(c("peakgroup_mz", "peakgroup_rt", "match_ID2"))
ms2_dt3 <- ms2_features_pos_core %>% dplyr::rename(id_pos_core = id)

#join with sql allowing multiple matches
id3 <- sqldf("select *
from peaklist_dt3 peak
left join ms2_dt3 ms2 on abs(ms2.ms2_precursor - peak.peakgroup_mz) < 0.05
	and abs(ms2.ms2_rt - peak.peakgroup_rt) < 15.0")

#filter ms2 id table to only include unique ms2 ids per xcms peakgroup 
id3_filt <- id3 %>% filter(!is.na(ms2_precursor) & !is.na(id_pos_core)) %>% select(-c("match_ID2")) %>% unique() %>% select(-c(peakgroup_rt, ms2_rt, ms2_precursor, ms2_feature_id)) %>% dplyr::rename(mtc_pos_core_frag = mtc)

#rejoin with peaklist
peaklist_hg_backgroundremoved_match_neg_poscore <- left_join(peaklist_hg_backgroundremoved_match_neg, id3_filt, by = "peakgroup_mz") 

#manually annotate 
peaklist_for_annotate <- peaklist_hg_backgroundremoved_match %>% select(c(compound_name, LOBdbase_mz, LOBdbase_ppm_match, peakgroup_mz, peakgroup_rt, xcms_peakgroup, ms2_precursor, ms2_rt, id_hg))
write.csv(peaklist_for_annotate, "output/annotation/fluids_woBA1A_for_annotation.csv")
```


# Upload Manually Curated Peaklist and Join with Full Peaklist
```{r eval = FALSE}
#import manually annotated peaklist
annotations <- readxl::read_excel("output/annotation/fluids_woBA1A_annotated_KRR_1_24_21.xlsx")
#select necessary compounds for joining
annotations <- annotations %>% filter(manually_identified == "yes")
annotations$id_hg <- as.factor(annotations$id_hg)

peaklist_annotated <- merge(peaklist_hg_backgroundremoved_match, annotations, by.y = c("compound_name", "LOBdbase_mz", "peakgroup_mz", "LOBdbase_ppm_match", "peakgroup_rt", "xcms_peakgroup", "ms2_precursor", "ms2_rt", "id_hg")) 

peaklist_annotated$ms2_feature_id<- recode(peaklist_annotated$ms2_feature_id, "FT09778" = "FT09766", "FT08244" = "FT08225")
annotated_list <- peaklist_annotated$ms2_feature_id

write.csv(peaklist_annotated, "output/annotation/fluids_woBA1A_annotated_peaklist.csv")
```


# Manually Check Identified Peaks
```{r eval = FALSE}
#table of identified compounds for plotting 
peaklist_annotated <- read.csv("output/annotation/fluids_woBA1A_annotated_peaklist.csv")
identified_compounds <- peaklist_annotated %>% select(c("compound_name", "peakgroup_mz", "peakgroup_rt")) 
#rename columns to work with function that requires columns named mz and rt 
identified_compounds <- identified_compounds %>% dplyr::rename(mz_db = peakgroup_mz, rt_db = "peakgroup_rt")
 
#define colors for sample groups
group_colors <- paste0(brewer.pal(3, "Dark2")[1:2], "60")
names(group_colors)<- c("blank", "sample")

#function to plot raw EICs
raw_XIC_plot <- function(db, raw) {
  group_colors <- paste0(RColorBrewer::brewer.pal(5, "Dark2"))
  names(group_colors)<- c("blank", "sample")
  
  for (row in 1:nrow(db)){
    XIC <- xcms::chromatogram(raw, mz = c(db$mz_db[row]-.005, db$mz_db[row]+.005), rt = c(db$rt_db[row] - 240, db$rt_db[row] + 240)) 
   plot(XIC, col = group_colors[raw$sample_group], main = db$compound_name[row])
    legend("topright", legend = c("blank", "sample"), fill = group_colors)
  }
}

#plot
raw_XIC_plot(db = identified_compounds, raw = dda_data_fluids_woBA1A_pos)
```

# Plot spectra from identified features
```{r eval = FALSE}
plot_spectra_annotate(spectra = dda_combined_fluids_woBA1A_pos_pp1, fragment_db = fragments_pos, nl_db = nl_pos, background_db = background_pos, to_include = annotated_list)
```

